Gammapy: A Python package for gamma-ray astronomy

### 2023-05-24 src/ms-review.tex
Partial rewrite of Section 4.6

\subsection{Surveys, Catalogs, and Population Studies}
\label{ssec:surveys-catalogs-and-population-studies}

Sky surveys have a large potential for new source detections, and discovery of new phenomena
 in \gammaray astronomy. They also offer less selection bias to perform
source population studies over a large set of coherently detected and modelled objects.
Early versions of \gammapy were developed in parallel to the preparation of
the \hess Galactic plane survey catalog~\citep[HGPS, ][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity and resolution provided by the new generation of
instruments scales up the number of detectable sources and the complexity of 
models needed to represent them accurately. As an example, if we compare the
results of the HGPS to the expectations from the \cta Galactic Plane survey
simulations, we jump from 78 sources detected by \hess to about 500 detectable by
CTA~\citep{Remy2021}. This large increase in the amount of data to analyse
and increase in complexity of modelling scenarios, requires the high-level
analysis software to be both scalabale as well as performant.

In short, the production of catalogs from \gammaray surveys can be divided in
four main steps: data reduction; object detection; model fitting and model
selection; associations and classification. All steps can either be done directly
with \gammapy or by relying on the seamless integration
of \gammapy with the scientific Python ecosystem. This allows to rely
on 3rd party functionality wherever needed.

The \iacts data reduction step is done in the same way described in the 
previous sections but scaled up to few thousands of observations. The object
detection step typically consists in finding local maxima in the significance or TS maps,
computed by the \code{ExcessMapEstimator} or \code{TSMapEstimator} respectively.
 Further refinements can
include for example filtering and detection on these maps with techniques from
the \textit{Scikit-image} package~\citep{scikit-image}, and outlier detection from
the \textit{Scikit-learn} package~\citep{scikit-learn}. This allows e.g., to 
reduce the number of spurious detections at this stage using standard
classification algorithms and then speed up the next step,
as less objects will have to be fitted simultaneously. During the modelling
step each object is alternatively fitted with different models in order to
determine their optimal parameters, and the best-candidate model. The
subpackage \code{gammapy.modeling.models} offers a large variety of choices, and the
possibility to add custom models.  Several spatial models (point-source, disk,
Gaussian...), and spectral models (power law, log parabola...) may be tested
for each object, so the complexity of the problem increases rapidly in regions
crowded with multiple extended sources. Finally an object is discarded if its
best-fit model is not significantly preferred over the null hypothesis (no
source) comparing the difference in log likelihood between these two
hypotheses.

For the association and classification step, which is tightly connected to the
population studies, we can easily compare the fitted models to the set of
existing \gammaray catalogs available in \code{gammapy.catalog}. Further
multi-wavelength cross-matches are usually required to characterize the
sources. This can easily be achieved by relying on coordinate
handling from Astropy in combination with affiliated packages
\textit{Astroquery}~\citep{astroquery}.

Studies performed on simulations not only offer a first glimpse on what could
be the sky seen by CTA (according to our current knowledge on source
populations), but also give us the opportunity to test the software on complex
use cases\footnote{Note that the CTA-GPS simulations were performed with the
	\textit{ctools} package~\citep{2016A&A...593A...1K} and analysed with both
	\textit{ctools} and \textit{gammapy} packages in order to cross-validate
	them.}. In this way we can improve performance, optimize our analyse strategies, and
identify the needs in terms of parallelisation to process the large datasets
provided by surveys.

>>>>

\subsection{Surveys, Catalogs, and Population Studies}
\label{ssec:surveys-catalogs-and-population-studies}
As a last application example we describe the use of \gammapy for large scale
analyses such as \gammaray surveys, catalogs and population studies.
Early versions of \gammapy were developed in parallel to the preparation of
the \hess Galactic plane survey catalog~\citep[HGPS, ][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,2018A&A...612A...3H}.

The increase in sensitivity and angular resolution provided by the new generation of
instruments scales up the number of detectable sources and the complexity of 
models needed to represent them accurately. As an example, if we compare the
results of the HGPS to the expectations from the \cta Galactic Plane survey
simulations, we jump from 78 sources detected by \hess to about 500 detectable by
CTA~\citep{Remy2021}. This large increase in the amount of data to analyse
and increase in complexity of modelling scenarios, requires the high-level
analysis software to be both scalabale as well as performant.

In short, the production of catalogs from \gammaray surveys can be divided in
four main steps: (a) data reduction, (b) object detection, (c) model fitting and model
selection and (d) associations and classification. All steps can either be done directly
with \gammapy or by relying on the seamless integration of \gammapy with the
scientific Python ecosystem. This allows to rely on 3rd party functionality
wherever needed. A simplifed catalog analysis based on \gammapy typically includes:

\begin{itemize}
	\item[a] The \iacts data reduction step is done in the same way described in the 
	previous sections but scaled up to few thousands of observations. This step can be
	trivially parallelised by deploying the \gammapy package on a cluster.
	\item[b] The object detection step typically consists in finding local maxima
	in the significance or TS maps, computed by the \code{ExcessMapEstimator} or
	\code{TSMapEstimator} respectively. For this \gammapy provides a simple \code{find_peaks}
	method. For more advanced methods users can rely on 3rd party packages such as
	\textit{Scikit-image}~\citep{scikit-image}. This packages provide e.g. general
	blob detection algorithms and image segmentation methods such as hysteresis thresholding
	or the watershed transform. 	
	\item[c] During the modelling step each object is alternatively fitted with different
	models in order to determine their optimal parameters, and the best-candidate model. The
	subpackage \code{gammapy.modeling.models} offers a large variety of choices for spatial
	and spectral models as well as the possibility to add custom models. For the model selection
	\gammapy provides statisticl helper methods to perform likelihood ratio tests.
	But users can also rely on 3rd party packages such as \textit{Scikit-learn}~\citep{scikit-learn}
	to compute quantities such as the Akaike information criterion (AIC) or the Bayesian
	information criterion (BIC), which also allow for model selection.
	\item[d] For the association and classification step, which is tightly connected to
	population studies, we can easily compare the fitted models to the set of
	existing \gammaray catalogs available in \code{gammapy.catalog}. Further
	multi-wavelength cross-matches are usually required to characterize the
	sources. This can easily be achieved by relying on coordinate
	handling from Astropy in combination with affiliated packages such as
	\textit{Astroquery}~\citep{astroquery}. For more advanced source classification
	methods users can again rely e.g. \textit{Scikit-learn} to perform supervised
	or unsupervised clustering.
\end{itemize}

\gammapy has been succesfully used for catalog studies performed on simulations
of the future \cta Galactic Plane Survey~\citep{Remy2021}. Besides the scientific
insights they also gave us the opportunity to test the \gammapy software on complex
use cases. This resulted in a number of improvements to the \gammapy package, such as
as improved performance, optimized analysis strategies documented to users, and
identifying the needs for future developments, including solutions for distributed
computing. The catalog studies of the \cta Galactic Plane Survey simulations
also allowed for detailed cross-comparison of the results obtained with the independent
\gammaray analysis package \textit{ctools}\citep{2016A&A...593A...1K}, with very consistent 
results.
